---
title: "Assingment NYC Flight Data"
author: "Nimish Kulkarni"
date: "October 10, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, `airports.csv`, `planes.csv` and `weather.csv`.  Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

you can set the working directory running the command
setwd("Path-to/CSX460/02-fundamentals")

```{r}
library(dplyr)
library(readr)

flightsData <- read_delim("data/flights.csv",delim = ",")
```

flightsData contains 336776 observations and 19 variables
## Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}
hist(flightsData$arr_delay, breaks=30, col="red")
hist(flightsData$dep_delay, breaks=50, col="green")
```

## Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}
plot(flightsData$day, flightsData$arr_delay, main="Date vs Arrival Delay",
   xlab="Date ", ylab="Arrival Delay ", pch=19) 
```



## Create tables for Categorical Variables

Create a `table` of occurences of origin airports and carrier.

```{r}


```


## Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}
airportsData <- read_delim("data/airports.csv",delim = ",")
planesData <- read_delim("data/planes.csv",delim = ",")
weatherData <- read_delim("data/weather.csv",delim = ",")
```



