---
title: "NYCFlights: Arrival Delay Regression Model"
author: "Nimish Kulkarni"
date: "Oct 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(ggplot2)
library(magrittr)
library(lubridate)
library(stringr)
library(dplyr)
library(data.table)
library(mlbench)
library(caret)
library(sqldf)
```

# NYCFlights Model

Using the rectangular data that you created last week and following the *Predicting Medical Expenses* example from the text (*MLwR*), create a model for arr_delay. Follow MLwR structure 
for building a model. Describe/Explain each of the steps and show all work in codeblocks below.


## Step 1: Collect Data 

```{r "Step 1: Collect Data" }
# Data Frames

flights  <- "../data/flights.csv" %>% read_csv 
planes   <- read_csv("../data/planes.csv")
airports <- read_csv("../data/airports.csv") 
weather  <- read_csv("../data/weather.csv")


# Merging Data Frames
YX <- flights 

## planes
YX <- merge( YX, planes, by="tailnum", all.x = TRUE )

## airports  
YX <- merge(YX, airports, by.x="origin", by.y="faa", all.x=TRUE, suffixes=c("", ".origin")  )
YX <- merge(YX, airports, by.x="dest", by.y="faa", all.x=TRUE, suffixes=c("", ".dest")  )

## weather
# weather$date <- ymd(weather$year, weather$month, weather$day)
YX <- 
  merge( flights, weather, by=c("origin","year","month","day","hour"), all.x=TRUE, suffixes=c('',".weather") )

## clearing the console screen
cat("\014")

```

## Step 2: Explore and Prep The Data


One of the things not done in the MLwR text is a pairwise comparison between the response and each of the predictors. Make sure to do this; this is often very illustrative of the relationship between that predictor and the response. This can be done with `pairs` or `psych::panel.pairs`


```{r "Step 2: Explore and Prep The Data" }
str(YX)
colnames(YX)
summary(YX$arr_delay)
hist(log(YX$arr_delay))
YX[c("year","month","day","hour","dep_time","sched_dep_time","dep_delay","arr_time","sched_arr_time","arr_delay","air_time","distance","minute","temp", "dewp", "humid","wind_dir","wind_speed","wind_gust","precip","pressure","visib")][is.na(
  YX[c("year","month","day","hour","dep_time","sched_dep_time","dep_delay","arr_time","sched_arr_time","arr_delay","air_time","distance","minute","temp", "dewp", "humid","wind_dir","wind_speed","wind_gust","precip","pressure","visib")]
)] <- 0

cor(YX[c("dep_time","sched_dep_time","dep_delay","arr_time","sched_arr_time","arr_delay","air_time","distance","temp","dewp","humid","wind_dir","wind_speed","wind_gust","precip","pressure","visib")])

# removing arr_delay to see which features are highly correlated
correlationMatrix = cor(YX[c("dep_time","sched_dep_time","dep_delay","arr_time","sched_arr_time","air_time","distance","temp","dewp","humid","wind_dir","wind_speed","wind_gust","precip","pressure","visib")])
print(correlationMatrix)

```

## Step 3:  Train The Model

```{r "Step 3: Train The Model" }

# train the model
model1 <- lm(YX$arr_delay~ YX$temp + YX$dewp + YX$humid + YX$dep_time, data=YX)
model2 <- lm(YX$arr_delay~ YX$dep_time + YX$sched_dep_time + YX$dep_delay + YX$arr_time+ YX$sched_arr_time + YX$air_time + YX$distance+ YX$temp + YX$wind_dir + YX$wind_speed + YX$wind_gust + YX$precip + YX$pressure + YX$visib + YX$dewp + YX$humid , data=YX)
model3 <- lm(YX$arr_delay~ YX$dep_time + YX$sched_dep_time + YX$dep_delay + YX$arr_time + YX$air_time + YX$distance+ YX$temp + YX$wind_dir + YX$precip + YX$pressure + YX$visib + YX$dewp, data=YX)
```

## Step 4: Evaluate Performance

Evaluating performance is a crutial step, that has yet to be treated well by the MLwR. Think about some of the tools that you have for evaluating performance.  Choose one and articulate why you have chosen it.

```{r "Step 4: Evaluate Performance" }
summary.lm(model1)
summary.lm(model2)
summary.lm(model3)
```


## Step 5: Improve Performance 

Show some steps for improving model performance.

```{r "Step 5: Improve Performance" }

```


# Question:

Is this a good model?  (Write your answer here.)

It seems to be a good initial model since we got R squared as 0.8456
