---
title: "NYCFlights: Arrival Delay Logistic Model"
author: "Nimish Kulkarni"
date: "23 October 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dtplyr)
library(readr)
library(ggplot2)
library(magrittr)
library(lubridate)
library(stringr)
library(plotly)
library(knitr)
library(dplyr)
library(data.table)
library(mlbench)
library(caret)
library(sqldf)
library(gplots)
library(ROCR)

```


## Logsitic and Inverse Logistic Transformation

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.
Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`
Plot the Inverse Logistic function from `[0,1]`

**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`

```{r "Logistic and Inverse Logistic" }
logistic.func <- function(x) {
  1 / (1 + exp(-x))
}
plot(logistic.func(-10:10))

inv.logistic.func <- function(x) {
  log(x/(1-x))
}
logistic.func(10)
inv.logistic.func(0.9999546)
plot(inv.logistic.func(0.1:0.9))
```


# NYCFlights Model

Using the rectangular data that you created from assignment and following the *Predicting Medical Expenses* example from the text (*MLwR*), create a model for arr_delay >= 15 minutes. Follow *MLwR* structure   for building a model. Describe/Explain each of the steps and show all work in codeblocks below.

KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file.   

## Step 1: Collect Data 

```{r "Step 1: Collect Data" }
# Data Frames

flights  <- "../data/flights.csv" %>% read_csv 
planes   <- read_csv("../data/planes.csv")
airports <- read_csv("../data/airports.csv") 
weather  <- read_csv("../data/weather.csv")


# Merging Data Frames
YX <- flights 

## planes
YX <- merge( YX, planes, by="tailnum", all.x = TRUE, suffixes=c("", ".pl") )

## airports  
YX <- merge(YX, airports, by.x="origin", by.y="faa", all.x=TRUE, suffixes=c("", ".origin")  )
YX <- merge(YX, airports, by.x="dest", by.y="faa", all.x=TRUE, suffixes=c("", ".dest")  )

## weather
# weather$date <- ymd(weather$year, weather$month, weather$day)
YX <- 
  merge( YX, weather, by=c("origin","year","month","day","hour"), all.x=TRUE, suffixes=c('',".weather") )

## clearing the console screen
cat("\014")
```

## Step 2: Explore and Prep The Data


One of the things not done in the MLwR text is a pairwise comparison between the response and each of the predictors. Make sure to do this; this is often very illustrative of the relationship between that predictor and the response. This can be done with `pairs` or `psych::panel.pairs`


```{r "Step 2: Explore and Prep The Data" }
yx <- copy(YX)

#copying arr_delay as new column to perform transformations
yx$arr_delay_new <- yx$arr_delay

summary(yx$arr_delay_new)

# handling NA values , thinking to replace with mean but since mean is less than 15 minutes
# assigning 0 class
yx["arr_delay_new"][is.na(yx["arr_delay_new"])] <- 0

# if arr_delay < 15 minutes then 0
yx["arr_delay_new"][yx["arr_delay_new"] < 15] <- 0

# assigning classes if arr_delay >= 15 minutes then 1
yx["arr_delay_new"][yx["arr_delay_new"] >= 15] <- 1



# commenting test code
# count = 0
# for (val in yx$arr_delay_new) {
#     if(val >= 15)  {
#       print(val)
#       count = count+1
#     }
# }
# print(count)


```

## Step 3:  Train The Model

```{r "Step 3: Train The Model" }
logModel1 <- glm( yx$arr_delay_new ~ yx$temp + yx$dewp + yx$humid + yx$dep_time , data = yx, family = "binomial")


yx[c("year","month","day","hour","dep_time","sched_dep_time","dep_delay","arr_time","sched_arr_time","air_time","distance","minute","temp", "dewp", "humid","wind_dir","wind_speed","wind_gust","precip","pressure","visib")][is.na(
  yx[c("year","month","day","hour","dep_time","sched_dep_time","dep_delay","arr_time","sched_arr_time","air_time","distance","minute","temp", "dewp", "humid","wind_dir","wind_speed","wind_gust","precip","pressure","visib")]
)] <- 0

logModel2 <- glm( yx$arr_delay_new ~ yx$dep_time + yx$sched_dep_time + yx$dep_delay + yx$arr_time + yx$air_time + yx$distance+ yx$temp + yx$wind_dir + yx$precip + yx$pressure + yx$visib + yx$dewp)

# model for part B , preparing it here
logModel3 <- glm( yx$arr_delay_new ~ yx$dep_time + yx$sched_dep_time + yx$temp + yx$wind_dir + yx$precip + yx$pressure + yx$dewp)

```

## Step 4: Evaluate Performance

Think about some of the tools that you have for evaluating performance.  Choose one and articulate why you have chosen it.

```{r "Step 4: Evaluate Performance" }
summary(logModel1)
summary(logModel2)
#confint(logModel1)
#confint(logModel2)

predictions <- predict(logModel1, newdata = yx, type="response")
labels <- yx$arr_delay_new

require(ROCR)
# Prepare data for plotting

pred <- with(yx, prediction(predictions, labels))
perf <- performance(pred, measure="acc", x.measure="cutoff")

# Get the cutoff for the best accuracy
bestAccInd <- which.max(perf@"y.values"[[1]])
bestMsg <- paste("best accuracy=", perf@"y.values"[[1]][bestAccInd], 
              " at cutoff=", round(perf@"x.values"[[1]][bestAccInd], 4))

plot(perf, sub=bestMsg)

# deleting the labels from data
yx <- subset(yx, select = -c(arr_delay_new) )


predictions <- predict(logModel1, newdata = yx, type="response")
predictions <- round(predictions)
mean(predictions==labels)

predictions <- predict(logModel2, newdata = yx, type="response")
predictions <- round(predictions)
mean(predictions==labels)

```

 
## Step 5: Improve Performance 

Show some steps for improving model performance.

```{r "Step 5: Improve Performance" }
# Model 2 has the highest accuracy of 90% around threshold of 0.3
library(ggplot2)
threshold <- seq(from=0.1,to=0.9,by=0.01)
j <- 1
accuracy <-c()
for(i in threshold){
  predictions <- predict(logModel2,newdata=yx,type = 'response')
  predictions <- ifelse(predictions>i,1,0)
  accuracy[j] <- mean(predictions==labels)
  j=j+1
  
}
model_accuracy <- data.frame(threshold,accuracy)
(ggplot(model_accuracy,aes(x=threshold,y = accuracy))+geom_line()+geom_point()+ggtitle("Threshold vs Accuracy"))

kable(model_accuracy %>% arrange(desc(accuracy)))

# rounding.func <- function(x) {
#   if (x >= 0.3){
#         return (1)
#   }
#   else
#         return (0)
# }


# using threshold value
predictions <- predict(logModel2, newdata = yx, type="response")
predictions[predictions < 0.32] <- 0
predictions[predictions >= 0.32] <- 1
mean(predictions==labels)    
```


# Question:

Is this a good model?  (Write your answer here.)

Yes, I believe this model2 is decent for basic predictions. Around 0.32 threshold we get 90 % Accuracy.

# PART B:

Your model should be good at explaining tardiness. Now, assume that your job is to predict arrival delays a month in advance. You can no longer use all the features in your model. Retrain your model using only features that will be *known* only a month in advance of the departure time.  Show all steps as above.

```{r "PART B" }

# all the steps are as above except building the model.
# I have removed the variables which I believe cannot be predicted before a month.

summary(logModel3)
#confint(logModel3)

predictions <- predict(logModel3, newdata = yx, type="response")
predictions <- round(predictions)
mean(predictions==labels)
```
