---
title: "Improving Model Perfromance / Tuning Parameters"
author: "Nimish Kulkarni"
date: "Dec 04, 2016"
output: html_document
---


## Tuning Parameter

Generically and regardless of model type, what are the purposes of a model
tuning parameters?

Often but not always control the variance-bias trade off

```
```

## Caret Models

This assignment demonstrates the use of caret for constructing models. Each
model should be built and compared using using `Kappa` as the performance
metric calculated using 10-fold repeated cross-validation with 3 folds.

Using the rectangular data that you created for the NYCFlights to create a model
for arr_delay >= 15 minutes.

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(dtplyr)
library(readr)
library(ggplot2)
library(magrittr)
library(lubridate)
library(stringr)
library(plotly)
library(knitr)
library(dplyr)
library(data.table)
library(mlbench)
library(caret)
library(sqldf)
library(gplots)
library(ROCR)
library(pROC)
library(randomForest)
```

```{r setup, include=FALSE}
#flights  <- "../data/flights.csv" %>% read_csv 
flights  <- read_csv("../data/flights.csv")
planes   <- read_csv("../data/planes.csv")
airports <- read_csv("../data/airports.csv") 
weather  <- read_csv("../data/weather.csv")


# Merging Data Frames
YX <- flights 

## planes
YX <- merge( YX, planes, by="tailnum", all.x = TRUE, suffixes=c("", ".pl") )

## airports  
YX <- merge(YX, airports, by.x="origin", by.y="faa", all.x=TRUE, suffixes=c("", ".origin")  )
YX <- merge(YX, airports, by.x="dest", by.y="faa", all.x=TRUE, suffixes=c("", ".dest")  )

## weather
# weather$date <- ymd(weather$year, weather$month, weather$day)
YX <- 
  merge( YX, weather, by=c("origin","year","month","day","hour"), all.x=TRUE, suffixes=c('',".weather") )

## clearing the console screen
cat("\014")
```

```{r}
# prep data and modify for classification

yx <- copy(YX)
#yx["arr_delay"][is.na(yx["arr_delay"])] <- 0

# if arr_delay < 15 minutes then 0
yx["arr_delay"][yx["arr_delay"] < 15] <- 0

# assigning classes if arr_delay >= 15 minutes then 1
yx["arr_delay"][yx["arr_delay"] >= 15] <- 1

yx <- yx[names(yx)[1:49]]
yx <- data.frame(lapply(yx, as.character), stringsAsFactors=FALSE)
yx <- data.frame(lapply(yx, as.numeric))
yx[is.na(yx)] <- 0
yx$arr_delay <- as.factor(yx$arr_delay)

set.seed(998)
inTraining <- createDataPartition(yx$arr_delay, p = .7, list = FALSE)
training <- yx[ inTraining,]
testing  <- yx[-inTraining,]

feature.names=names(training)

for (f in feature.names) {
  if (class(training[[f]])=="factor") {
    levels <- unique(c(training[[f]]))
    training[[f]] <- factor(training[[f]],
                   labels=make.names(levels))
  }
}

feature.names=names(testing)

for (f in feature.names) {
  if (class(testing[[f]])=="factor") {
    levels <- unique(c(testing[[f]]))
    testing[[f]] <- factor(testing[[f]],
                   labels=make.names(levels))
  }
}
```

- glm
- rpart
- knn
- C50
- randomForest
- adaBoost
- Two methods of your choice from the Caret Model List (you will need to install any dependencies)

Save the caret objects with the names provided.
```{r}

myControl <- trainControl(method='cv', number=3, repeats=10, 
                          returnResamp='none', classProbs=TRUE,
                          returnData=FALSE, savePredictions=TRUE, 
                          verboseIter=TRUE, allowParallel=TRUE,
                          summaryFunction=twoClassSummary)

```

```{r}

# Your work here.

fit.glm <- train(arr_delay~., 
                  data = training,
                  method='glm', 
                  trControl=myControl,  
                  metric = "ROC",
                  na.action = na.pass,
                  preProc = c("center", "scale"))
fit.knn <- train(arr_delay~., 
                  data = training,
                  method='knn', 
                  trControl=myControl,  
                  metric = "Kappa",
                  na.action = na.pass,
                  preProc = c("range"))
fit.rpart <- train(arr_delay~., 
                  data = training,
                  method='rpart', 
                  trControl=myControl,  
                  metric = "ROC",
                  na.action = na.pass,
                  preProc = c("center", "scale"))
fit.rf <- train(arr_delay~., 
                  data = training,
                  method='rf', 
                  trControl=myControl,
                  metric = "ROC",
                  na.action = na.pass,
                  preProc = c("center", "scale"))

#fit.myown1 <- ..
#fit.myown1 <- ..
```

Compare the  models?

Which is best?  Why?
```{r}
summary(fit.glm)
summary(fit.knn)
summary(fit.rpart)
summary(fit.rf)
```
```
